### 2020-09-15 Manuscript/manuscript.tex
progress on glossary

\documentclass[11pt,letterpaper]{article}
\setcounter{section}{3}
\usepackage{threeparttable}
\usepackage{textcomp,marvosym}
\usepackage{amsmath,amssymb}
\usepackage[left]{lineno}
\usepackage{changepage}
\usepackage{rotating}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{pdfpages}
\usepackage{url}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage[outercaption]{sidecap}
\usepackage{glossaries}
\doublespacing

@@

\makeglossaries

\newglossaryentry{latex}
{
    name=latex,
    description={Is a mark up language specially suited 
    for scientific documents}
}

\newglossaryentry{Laurentia}
{
    name=Laurentia,
    description={craton that forms the Precambrian core of North America}
}

\newglossaryentry{Nuna}
{
    name=Nuna,
    description={hypothesized Paleoproterozoic to Mesoproterozoic supercontinent}
}

\newglossaryentry{Archean}
{
    name=Archean,
    description={Geologic eon spanning from 4,000 to 2,500 million years ago (4 to 2.5 Ga).}
}

@@

Laurentia refers to the craton that forms the Precambrian interior of North America and Greenland (Fig. \ref{fig:Laurentia_map}). Laurentia comprises multiple \gls{Archean} provinces that had unique histories prior to their amalgamation in the Paleoproterozoic ca. 1.8 Ga, as well as tectonic zones of crustal growth that post-date this assembly \citep{Hoffman1989a, Whitmeyer2007a}. Collision between the Superior province and the composite Slave+Rae+Hearne provinces that resulted in the Trans-Hudson orogeny represents a major event in the formation of Laurentia \citep{Corrigan2009a}. Terminal collision recorded in the Trans-Hudson orogen is estimated to have been ca. 1.86 to 1.82 Ga based on constraints such as U-Pb dating of monazite grains and zircon rims \citep[e.g.]{Skipton2016a, Weller2017a}. A period of accretionary and collisional orogenesis is recorded in the constituent provinces and terranes of Laurentia leading up to the terminal collision of the Trans-Hudson orogeny. This overall story of rapid Paleoproterozoic amalgamation of Laurentia's constituent Archean provinces, including the terminal Trans-Hudson orogeny, was synthesized in the seminal \textit{United Plates of America} paper of \citet{Hoffman1988a} and has been refined in the time since -- particularly with additional geochronological constraints. Of most relevance here are the events that led to the suturing of more major Archean provinces: the Thelon orogen associated with the collision between the Slave and Rae provinces ca. 2.0 to 1.9 Ga \citep{Hoffman1989a}; the Snowbird orogen associated with ca. 1.89 Ga collision between the Rae and Hearne provinces and associated terranes \citep{Berman2007a}; the Nagssugtoqidian orogen due to the ca. 1.86 to 1.84 Ga collision between the Rae and North Atlantic provinces \citep{St-Onge2009a}; and the Torngat orogen resulting from the ca. 1.87 to 1.85 Ga collision of the Meta Incognita province (grouped with the Rae province in older compilations) with the North Atlantic province \citep{St-Onge2009a}.

As for the suturing of the Wyoming province to Laurentia, many models posit that it was conjoined with Hearne and associated provinces at the time of the Trans-Hudson orogeny \citep[e.g.][]{St-Onge2009a, Pehrsson2015a} or was proximal to Hearne and Superior while still undergoing continued translation up to ca. 1.80 Ga \citep{Whitmeyer2007a}. A contrasting view has been been proposed that the Wyoming province and Medicine Hat blocks were not conjoined with the other Laurentia provinces until ca. 1.72 Ga \citep{Kilian2016b}. This interpretation is argued to be consistent with geochronological constraints on monazite and metamorphic zircon indicating active collisional orogenesis associated with the Big Sky orogen on the northern margin of the craton as late as ca. 1.75 to 1.72 Ga \citep{Condit2015a} and ca. 1.72 tectonomagmatic activity in the Black Hills region \citep{Redden1990a}. However, the evidence for earlier orogenesis ca. 1.78 to 1.75 in the Black Hills \citep{Dahl1999a,Hrncir2017a}, as well as high-grade metamorphism as early as ca. 1.81 Ga in the Big Sky orogen \citep{Condit2015a}, may support the interpretation of \citet{Hrncir2017a} that ca. 1.72 Ga activity is a minor overprint on ca. 1.75 terminal suturing between Wyoming and Superior. Regardless, in both of these interpretations, Wyoming is a later addition to Laurentia with final suturing post-dating ca. 1.82 Ga amalgamation of Archean provinces with the Trans-Hudson orogen further to the northeast. Overall, the collision of these Archean microcontinents between ca. 1.9 and 1.8 Ga led to rapid amalgamation of the majority of the Laurentia craton (Fig. \ref{fig:Laurentia_map}).

@@

\clearpage
\printglossaries

>>>>

\documentclass[11pt,letterpaper]{article}
\setcounter{section}{3}
\usepackage{threeparttable}
\usepackage{textcomp,marvosym}
\usepackage{amsmath,amssymb}
\usepackage[left]{lineno}
\usepackage{changepage}
\usepackage{rotating}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{pdfpages}
\usepackage{url}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage[outercaption]{sidecap}
\doublespacing

@@

@@

Laurentia refers to the craton that forms the Precambrian interior of North America and Greenland (Fig. \ref{fig:Laurentia_map}). Laurentia comprises multiple Archean provinces that had unique histories prior to their amalgamation in the Paleoproterozoic ca. 1.8 Ga, as well as tectonic zones of crustal growth that post-date this assembly \citep{Hoffman1989a, Whitmeyer2007a}. Collision between the Superior province and the composite Slave+Rae+Hearne provinces that resulted in the Trans-Hudson orogeny represents a major event in the formation of Laurentia \citep{Corrigan2009a}. Terminal collision recorded in the Trans-Hudson orogen is estimated to have been ca. 1.86 to 1.82 Ga based on constraints such as U-Pb dating of monazite grains and zircon rims \citep[e.g.]{Skipton2016a, Weller2017a}. A period of accretionary and collisional orogenesis is recorded in the constituent provinces and terranes of Laurentia leading up to the terminal collision of the Trans-Hudson orogeny. This overall story of rapid Paleoproterozoic amalgamation of Laurentia's constituent Archean provinces, including the terminal Trans-Hudson orogeny, was synthesized in the seminal \textit{United Plates of America} paper of \citet{Hoffman1988a} and has been refined in the time since -- particularly with additional geochronological constraints. Of most relevance here are the events that led to the suturing of more major Archean provinces: the Thelon orogen associated with the collision between the Slave and Rae provinces ca. 2.0 to 1.9 Ga \citep{Hoffman1989a}; the Snowbird orogen associated with ca. 1.89 Ga collision between the Rae and Hearne provinces and associated terranes \citep{Berman2007a}; the Nagssugtoqidian orogen due to the ca. 1.86 to 1.84 Ga collision between the Rae and North Atlantic provinces \citep{St-Onge2009a}; and the Torngat orogen resulting from the ca. 1.87 to 1.85 Ga collision of the Meta Incognita province (grouped with the Rae province in older compilations) with the North Atlantic province \citep{St-Onge2009a}.

As for the suturing of the Wyoming province to Laurentia, many models posit that it was conjoined with Hearne and associated provinces at the time of the Trans-Hudson orogeny \citep[e.g.][]{St-Onge2009a, Pehrsson2015a} or was proximal to Hearne and Superior while still undergoing continued translation up to ca. 1.80 Ga \citep{Whitmeyer2007a}. A contrasting view has been been proposed that the Wyoming and Medicine Hat provinces were not conjoined with the other Laurentia provinces until ca. 1.72 Ga \citep{Kilian2016b}. This interpretation is argued to be consistent with geochronological constraints on monazite and metamorphic zircon indicating active collisional orogenesis associated with the Big Sky orogen on the northern margin of the craton as late as ca. 1.75 to 1.72 Ga \citep{Condit2015a} and ca. 1.72 tectonomagmatic activity in the Black Hills region \citep{Redden1990a}. However, the evidence for earlier orogenesis ca. 1.78 to 1.75 in the Black Hills \citep{Dahl1999a, Hrncir2017a}, as well as high-grade metamorphism as early as ca. 1.81 Ga in the Big Sky orogen \citep{Condit2015a}, may support the interpretation of \citet{Hrncir2017a} that ca. 1.72 Ga activity is a minor overprint on ca. 1.75 terminal suturing between Wyoming and Superior. Regardless, in both of these interpretations, Wyoming is a later addition to Laurentia with final suturing post-dating ca. 1.82 Ga amalgamation of Archean provinces with the Trans-Hudson orogen further to the northeast. Overall, the collision of these Archean microcontinents between ca. 1.9 and 1.8 Ga led to rapid amalgamation of the majority of the Laurentia craton (Fig. \ref{fig:Laurentia_map}).

@@

\section*{Glossary}
\noindent\textbf{accretionary orogeny } Lithospheric deformation associated with the subduction of oceanic lithosphere and the addition of material from the downgoing plate such as island arcs.

\noindent\textbf{Archean } A geologic eon spanning from 4,000 to 2,500 million years ago (4 to 2.5 Ga)

\noindent\textbf{Archean province } A contiguous area of Archean continental lithosphere typically surrounded by Proterozoic orogens inferred to be suture zones (e.g. Superior province). \\

\noindent\textbf{collisional orogeny } Lithospheric deformation resulting from the collision of two significant provinces of continental lithosphere.

\noindent\textbf{Ga } Giga-annum, one billion (10$^9$) years. This term is used as an abbreviation for ``billions of years before present.''

\noindent\textbf{Laurentia } The cratonic core of the North America continent that formed through the amalgamation of Archean provinces in the Paleoproterozoic.

\noindent\textbf{lithosphere } The rigid outermost layer of the Earth that is broken into tectonic plates and responds to the emplacement of a load by flexural bending.

\noindent\textbf{Ma } Mega-annum, one million (10$^6$) years. This term is used as an abbreviation for ``millions of years before present.''

\noindent\textbf{Manikewan Ocean } An ocean basin interpreted to have existed between the  Slave+Rae+Hearne+North Atlantic provinces and Superior province that closed leading up to the Trans-Hudson orogeny.

\noindent\textbf{Mesoproterozoic } A geologic era spanning from 1,600 to 1,000 million years ago.

\noindent\textbf{monazite } A phosphate mineral (Ce,La,Nd,Th)(PO\textsubscript{4},SiO\textsubscript{4}) found as an accessory phase in metamorphic rocks that can be targeted by U-Pb geochronology to date metamorphic events.

\noindent\textbf{Neoproterozoic } A geologic era spanning from 1,000 to 541 million years ago.

\noindent\textbf{Nuna } A hypothesized supercontinent interpreted to have formed late in the Paleoproterozoic era and broke apart in the middle of the Mesoproterozoic.

\noindent\textbf{orogen } A region of lithosphere that has undergone deformation during a mountain-building event (an orogeny).

\noindent\textbf{Paleoproterozoic } A geologic era spanning from 2,500 to 1,600 million years ago.

\noindent\textbf{Precambrian } A commonly used informal term to refer to geologic time prior to the Cambrian Period that started 541 million years ago.

\noindent\textbf{province } A spatial entity with a shared geologic history. The term is used in this chapter to refer to Archean provinces that moved as independent cratonic blocks prior to Laurentia's amalgmation (e.g. Superior province). It is also used to refer to zones of crustal growth associated with orogens and the products of contemporaneous magmatic activity (large igneous provinces).

\noindent\textbf{Rodinia } A hypothesized supercontinent interpreted to have formed late in the Mesoproterozoic era and broke apart in the middle of the Neoproterozoic.

\noindent\textbf{Superior province } The largest Archean province of Laurentia framed by the Trans-Hudson orogen to the west, the Grenvillian orogen to the east and the Penokean orogen to the south. It is also referred to as the Superior craton.

\noindent\textbf{Thelon orogen } An orogen resulting from a Paleoproterozoic collision orogeny between the Slave and Rae provinces.

\noindent\textbf{Wopmay orogen } An orogen resulting from collision between the Hottah terrane, a continental magmatic arc, and the west margin of the Slave province.

\noindent\textbf{zircon } A nesosilicate mineral with the chemical name of zirconium silicate and a chemical formula of ZrSiO\textsubscript{4}.
